---
title: "Estimating Access to Rabies Post-Exposure Prophylaxis in Madagascar"
author: ""
date: 
output:
  html_document: default
  pdf_document:
    includes:
      in_header: mystyle.sty
  word_document:
    reference_docx: format_docs/word_styles_reference_01.docx
css: style.css
---

## INTRODUCTION

Access to healthcare is a significant driver of disease burden globally. Spatial access to care (cite Jess paper here). Rarely incorporated into national estimates of burden due to lack of data and tools for understanding...

<br>

Canine rabies causes an estimated 59,000 human deaths annually. Post-exposure prophylaxis (PEP) prevents fatalities if delivered in a timely manner, however, access to this intervention is highly limited in areas where the disease is endemic. Data on true rabies exposures in humans and incidence in animals is also lacking in most of these countries. Often, the only data available are numbers of bite victims reporting to health facilities. The majority of rabies burden studies use these data to estimate burden of rabies from probability decision tree frameworks, often with the key assumption that overall reported bite incidence (i.e. both bites due to non-rabid and rabid animals) are proportional to rabies incidence (i.e. the more overall bites reported in a location, the higher the incidence of *rabies* exposures there) and that reporting is uniform across space. 

In Madagascar, Institut Pasteur de Madagascar (IPM) provides PEP at no-cost to the government or to patients that present to 31 clinics across the country. PEP is not available at any other public clinics or through the private sector. Given that rabies is endemic in the country and there is effectively no vaccination of domestic dogs, we use the baseline predictions of incidence from the dynamic model reported in Hampson et al. 2018. We flip the standard decision tree and make the assumption that reported bite incidence reflects access and reporting to PEP rather than differences in rabies incidence, using travel times to clinics as a predictor of bites reported to clinics. Assuming a range of exposure incidence given endemic rabies, we generate sub-national estimates of rabies burden and explore the impacts of expanding access to PEP in Madagascar. 

The most recent estimation of burden and the impact of PEP used another approach, using transmission dynamic models as the backbone to predict incidence based on level of vaccination coverage and size of the dog population at the national level. Using transmission dynamic models to estimate incidence could improve upon previous studies which may underestimate rabies burden in areas with low reporting. Here, we adapt this framework and use bite patient data from Madagascar to predict rabies burden.


## METHODS

### Bite patient data
We used the database of bite patient forms submitted from 24 ARMC across ARMC between 2014 - 2017. These were individual patient data forms that were submitted to IPM as frequently as monthly to annually by clinics. Two clinics, the IPM ARMC and the Fianarantsoa ARMC used computer databases from which the data during this period were extracted and merged to the larger database. These data include the administrative district of the bite patient's address. As previous analyses showed likely substantial undersubmission of forms from clinics (*cite baseline data paper*), we calculated the proportion of days of the year for which forms were submitted. We excluded periods with 10 or more consecutive dates with no patient forms and calculated the proportion of days included per year. 

- Patient data from 19 ARMC accross the country over four years (2013 - 2017)

  - includes dates and home locations (to the district level) of patient reporting for PEP
- Patient data from Moramanga ARMC over 28 months (Sep 2016 - Dec 2018)
  - includes dates and home locations (to the commune level) of patients reporting for PEP
 

RIG data?

<br>

### GIS Data
We use the global friction surface for 2015 generated by the Malaria Atlas Project ( https://map.ox.ac.uk/research-project/accessibility_to_cities/, Weiss et al. 2015,) and GPS points of clinics to get travel time to the nearest ARMC for the country of Madagascar at a 1 x 1 km scale. We then calculated a weighted average of travel times by human population to the commune level, using administrative shapefiles available trhough the UN Office for the Coordination of Humanitarian Affairs. For each clinic, we defined the catchment area as all **districts?** for which the clinic was the closest ARMC. Human population estimates were taken from the 2015 UN adjusted population projections from World Pop (www.worldpop.org, Linaird et al. 2012) and also aggreagated to the commune level
<br>

### Model of reported bites as a function of travel time
While the national bite patient data is available at the district level, travel times vary significantly within a district (see Figure SX). In order to translate the impacts of differences in access at sub-district scales to the magnitude of reported bites at the district scale, we modeled bites at the district level as the sum of incidence at the commune level. Incidence at the commune level is then a function of travel times to the closest ARMC. Specifically, we modeled bites as follows:
$$ 
  \mu_{d} = \sum \limits_{j=1}^jexp(\beta_{t}T_j + \beta_0)\times pop_j 
$$
where $\mu_{d}$ is the mean number of bites in district, which is the sum of bites at the commune level given **commune level travel times**, $T_j$. We then estimate the likelihood of observing the bites at the district level where bites are a poisson distribution around the mean $\mu_{d}$, given $\beta_t$, the effect of travel times of reported bites and $\beta_0$, the model intercept.

As we had data available on reported bites at the commune level from the Moramanga ARMC, we modeled observed commune bites in the same framework, but un-aggregated where:
$$ 
\mu_{j} = exp(\beta_tT_j + \beta_0)\times pop_j 
$$
where $\mu_{j}$ = the mean number of bites in commune $j$ and the observed bites at the commune level follow a poisson distribution around the mean $\mu_d$. We only included communes which were designated to be within the catchment for the clinic (so primarily the Moramanga and Anosibe A'nala Districts)?? We compared our estimates of $\beta_t$ (i.e. the impact of travel time on incidence) and $\beta_0$ (the intercept) for our district data at the national level and the commune level data from the Moramanga ARMC.

To better understand if ttimes might capture access, we also looked at whether pop alone can predict bites and distance could predict bites...
Ability to predict out-of-fit to other catchments? Tana + Fianarantsoa?

<br>

### Estimation of burden and reporting 
We used our model to predict average annual bite incidence for all 114 districts in Madagascar, and estimated deaths due to rabies given this and assumptions about rabies exposures.

We calculated deaths as:
$$
deaths_i = (R_i - p_{rabid}B_i) \times p_{death}
$$

where the proportion of reported bites that are rabies exposures ($p_{rabid}$) are defined as:
$$
           p_{rabid}=
            \begin{cases}
            x, & \text{if}\ \frac{R_i \times \rho_{max}}{B_i} > x \\
            \frac{R_i \times \rho_{max}}{B_i}, & \text{otherwise}
            \end{cases}
$$
<br>

such that rabid reported bites (i.e. $p_{rabid} \times B_i$) cannot exceed the expected number of human exposures given maximum reporting (i.e. $R_i \times \rho_{max}$). $\rho_{max}$ taken from the Moramanga ARMC data for Moramanga Ville, the commune with the ARMC (i.e. the area with the minimum travel time). We look at the range of $x$ reported in Rajeev et al. 2018 for data from the Moramanga District (0.2 - 0.6). 

$R_i$ is drawn from a uniform distribution between the minimum and maximum expected number of human exposures. We calculate this as:
    
<br>

$$r \times D \times p_{exp}$$ 
    
<br>

where $r$ is rabies incidence in dogs in the absence of any vaccination multiplied by the estimated dog population in the commune ($D_i$) and the exposure rate per rabid dogs ($p_{exp}$ = 0.39 persons exposed per rabid dog)(Hampson et al. 2018). We estimate the dog population by using a human:dog ratio of 5 to generate our maximum expected incidence and an HDR of 25 for our minimum expected incidence. As there is little data on dog populations in Madagascar, this range of HDRs encompasses a wide range observed across Africa (cite!). 

<br>

### Sensitivity Analyses
- R (increasing or decreasing with pop) (min + max)
- P-rabid min and max
- Rho-max min and max
- Probability of death min and max

<br>

### Estimating the impact of expanding PEP Access
We use this framework to compare three scenarios of PEP provisioning in Madagascar:

1. The baseline with the current clinic locations (n = 31)
2. Expansion of ARMC to one clinic per district (n = 114)
3. Expansion of ARMC to all CSB IIs, something to describe CSB IIs here, that is primary hospitals (n = X)

We use data on the location of **CSBs** provided by IPM to regenerate travel times to the nearest ARMC given expansion as per scenario 2 and 3. We then predict the expected bite incidence from the model given these new travel times and compare the relative decreases in burden for the three scenarios.
  
  
## RESULTS 

```{r directory, echo = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
rm(list = ls())

## Libraries
library(knitr)
library(RImagePalette)
library(jpeg)
library(tidyverse)
library(lubridate)
library(magrittr)
library(rgdal)
library(broom)
library(stringr)
library(ISOweek)
library(raster)
library(sp)

source("R/utils.R")

select <- dplyr::select
knitr::opts_chunk$set(fig.pos = 'h', echo = FALSE, message = FALSE, warning = FALSE, results = 'hide')
```


```{r get data}
## read in data
ctar_data <- read.csv("data/SaisieRage_DATA_2018-09-21_1755.csv")
ctar_metadata <- read.csv("data/ctar_metadata.csv")
mada_district <- readOGR("data/MadaGIS/MadaPops.shp")
mada_communes <- readOGR("output/communes/communes_extracted.shp")

## catchment data
dist_catchments <- read.csv("output/catchments_district_unmasked.csv", row.names = 1)
names(dist_catchments) <- c("mdg_dis_co", "CTAR", "ttimes_weighted")
comm_catchments <- read.csv("output/catchments_commune_unmasked.csv", row.names = 1)
names(comm_catchments) <- c("mdg_cm_", "CTAR", "ttimes_weighted")

## Centers with no data
no_data <- c("IPM", "Fianarantsoa", "Ambatomainty", "Ambovombe Androy",
             "Antsiranana", "Mandritsara", "Marolambo", "Nosy be", "Sainte Marie",
             "Taolagnaro", "Tsiroanomandidy", "Vangaindrano")
ctar_metadata$exclude <- 0
ctar_metadata$exclude[ctar_metadata$CTAR %in% no_data] <- 1

## Getting ctar district and commune
ctar_metadata$ctar_dist <- mada_district$mdg_dis_co[match(ctar_metadata$District, 
                                                           mada_district$district)]
pts <- SpatialPoints(cbind(ctar_metadata$LONGITUDE, ctar_metadata$LATITUDE), 
                   proj4string = CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))
ctar_metadata$ctar_comm <- over(pts, mada_communes)$mdg_cm_
dist_catchments$ctar_in_dist <- ifelse(dist_catchments$mdg_dis_co %in% ctar_metadata$ctar_dist, 1, 0)
comm_catchments$ctar_in_comm <- ifelse(comm_catchments$mdg_cm_ %in% ctar_metadata$ctar_comm, 1, 0)
```

### Clinic catchments and reporting
For most districts, the majority of bites were reported to the closest clinic as estimated by our travel time metric (Fig 1B), although some patients did report to the ARMC that were not closest to them in terms of travel times (Figure 1A). In addition, at the clinic level, the majority of bites reported to a given ARMC came from within the catchment (Figure 1C).

```{r network fig}
## Network fig
## Revise: points = CTAR points with size proportional to number of bites reported, lines to district proportional to the number of bites from that district, outline = by district, communes = polygons shaded by catchments?

## Get color palette from other network fig!
ctar_metadata$color <- c("#FCC56F","#FFDBE5", "#7A4900", "#CBCDD2", "#0000A6", "#EFF2F1",
                 "#99d8c9", "#B79762", "#004D43", "#8FB0FF", "#997D87", "#FD9C54", "#8362B5",
                 "#578FB0","#5A0007", "#809693", "#D16100", "#1B4400", "#4FC601", "#3B5DFF", 
                 "#4A3B53", "#FF2F80","#61615A", "#BA0900", "#6B7900", "#00C2A0", "#FFAA92",
                 "#FF90C9", "#B903AA", "#FEFFE6", "#E9E9D2")
ctar_metadata$color[ctar_metadata$CTAR %in% no_data] <- "#D3D3D3"
ctar_metadata$fill <- ctar_metadata$color
ctar_metadata$fill <- add.alpha(ctar_metadata$fill, 0.4)
display_palette(ctar_metadata$color)

dist_catchments %>% left_join(select(ctar_metadata, CTAR, exclude, 
                                     color, fill, ctar_dist, LATITUDE, LONGITUDE)) -> dist_catchments
mada_district@data <- left_join(mada_district@data, dist_catchments) ## need this?

## get exposure matrix = first set of points
## district exposures
ctar_data %>% 
  drop_na(id_ctar, district) %>%
  group_by(district) %>% # group by district
  summarize(n = n()) -> exps_dist # get count of exposures
## ctar exposures
ctar_data %>% 
  drop_na(id_ctar) %>%
  group_by(id_ctar) %>% # group by ctar
  summarize(n = n()) -> exps_ctar
## ctar by district exposure matrix
ctar_data %>% 
  drop_na(id_ctar, district) %>%
  group_by(district, id_ctar) %>% # group by district and ctar
  summarize(n = n()) -> exp_mat# get count of exposures

## To plot district bites
dist_points <- cbind(select(mada_district@data,
                            mdg_dis_co), coordinates(mada_district))
names(dist_points)[2:3] <- c("dist_long", "dist_lat")
dist_points %>%
 left_join(select(ctar_metadata, ctar_dist, LATITUDE, LONGITUDE), 
           by = c("mdg_dis_co" = "ctar_dist")) %>%
 mutate(dist_long = coalesce(LONGITUDE, dist_long), 
        dist_lat = coalesce(LATITUDE, dist_lat)) -> dist_points

exps_dist %>%
  right_join(select(dist_points, mdg_dis_co, dist_long, dist_lat),
             by = c("district" = "mdg_dis_co")) %>%
  left_join(dist_catchments, by = c("district" = "mdg_dis_co") ) -> dist_points
dist_points$size <- log(dist_points$n + 0.1)*0.25

## To plot ctar points (only those not excluded)
exps_ctar %>%
 left_join(select(ctar_metadata, id_ctar, LATITUDE, LONGITUDE, 
                  exclude, color, fill)) %>%
 filter(exclude == 0) -> ctar_points

## To plot lines between ctar + districts
exp_mat %>% 
  left_join(select(dist_points, district, dist_long, dist_lat)) %>%
  left_join(select(ctar_metadata, id_ctar, LATITUDE, LONGITUDE,
                   color, fill, ctar_dist)) -> dist_lines
dist_lines$size <- log(dist_lines$n + 0.1)*0.2

pdf("figs/network.pdf")
plot(mada_district, col = mada_district$fill, border = "grey") 
segments(dist_lines$dist_long, dist_lines$dist_lat, x1 = dist_lines$LONGITUDE, 
         y1 = dist_lines$LATITUDE,
         lwd = dist_lines$size*2, col = dist_lines$color)
points(ctar_points$LONGITUDE, ctar_points$LATITUDE, col = ctar_points$color, pch = 1,
       cex = log(ctar_points$n + 0.1)*0.25)
points(dist_points$dist_long, dist_points$dist_lat, cex = dist_points$size, pch = 20,
       col = dist_points$color)
dev.off()

exp_mat %>% 
  left_join(select(ctar_metadata, ctar_obs = CTAR, id_ctar)) %>%
  left_join(select(dist_catchments, mdg_dis_co, ctar_catch = CTAR, exclude), 
            by = c("district" = "mdg_dis_co")) %>%
  group_by(district) %>% # group by ctar
  mutate(prop = n/sum(n),
         catch_match = ifelse(ctar_obs == ctar_catch, 1, 0)) %>%
  filter(catch_match == 1) -> catch_dist
hist(catch_dist$prop, main = "", 
     xlab = "Proportion of patients \n reporting to closest ARMC",
     ylab = "Number of districts", col = "grey50", border = "white", xlim = c(0, 1))

exp_mat %>% 
  left_join(select(ctar_metadata, ctar_obs = CTAR, id_ctar)) %>%
  left_join(select(dist_catchments, mdg_dis_co, ctar_catch = CTAR, exclude), 
            by = c("district" = "mdg_dis_co")) %>%
  group_by(ctar_obs) %>% # group by ctar
  mutate(prop = n/sum(n),
         catch_match = ifelse(ctar_obs == ctar_catch, 1, 0)) %>%
  filter(catch_match == 1) %>%
  summarize(prop = sum(prop)) -> catch_ctar
hist(catch_ctar$prop, main = "", 
     xlab = "Proportion of patients reporting from within catchment",
     ylab = "Number of ARMC", col = "grey50", border = "white", xlim = c(0, 1))
```

We estimated clinic level reporting as the proportion of days on which forms were submitted, excluding any periods for which there were no forms submitted for 10 consecutive days (Figure 2). We further excluded any years for which there was less than 25% reporting. 

```{r clinic level reporting}
######### Patient time series
ctar_data %>%
  select(id_ctar, date_de_consultation) %>%
  drop_na(id_ctar) %>%
  mutate(date_de_consultation = ymd(date_de_consultation)) %>%
  gather(dose, date_de_consultation, -id_ctar) %>%
  group_by(id_ctar, date_de_consultation) %>%
  summarise(n = n()) -> patient_ts

patient_ts <- patient_ts[!is.na(patient_ts$date_de_consultation), ]
patient_ts$ctar <- ctar_metadata$CTAR[match(patient_ts$id_ctar, ctar_metadata$id_ctar)]
patient_ts <- patient_ts[!is.na(patient_ts$ctar), ]

start_date <- ymd("2014-01-01")
end_date <- ymd("2017-12-31")

## getting reporting
ts <- as_tibble(seq(start_date, end_date, by = "day"))

patient_ts %>%
  right_join(ts, by = c("date_de_consultation" = "value")) %>%
  select(id_ctar, date_de_consultation, n) %>%
  spread(id_ctar, n) %>%
  replace(., is.na(.), 0) -> doses_wide
dose_mat <- as.matrix(doses_wide[, 2:(ncol(doses_wide) - 1)])

get.days <- function(dmat = dose_mat, threshold = 10) {
  date_mat <- matrix(NA, nrow(dmat), ncol(dmat))
  for (j in 1:ncol(date_mat)){
    rle(dmat[ , j]) %>%
      unclass() %>%
      as.data.frame() %>%
      mutate(end = cumsum(lengths),
             start = c(1, lag(end)[-1] + 1)) %>%
      filter(values == 0, lengths >= threshold) -> rles
  if(nrow(rles) > 0){
    for (i in 1:nrow(rles)){
        date_mat[rles$start[i]:rles$end[i], j] <- 0
      }
    }
  }
  
  date_mat <- replace(date_mat, is.na(date_mat), 1)
  date_mat %>% 
    as_tibble() %>%
    group_by(date = year(doses_wide$date_de_consultation)) %>%
    summarise_all(funs(sum(.)/n())) -> clinic_reporting
  colnames(clinic_reporting) <- colnames(doses_wide)[1:(ncol(doses_wide)-1)]

  clinic_reporting <- gather(clinic_reporting, id_ctar, prop, -date_de_consultation)
  clinic_reporting$ctar <- ctar_metadata$CTAR[match(clinic_reporting$id_ctar, 
                                                  ctar_metadata$id_ctar)]
  clinic_reporting$id_ctar <- as.numeric(clinic_reporting$id_ctar)
  clinic_reporting$threshold <- threshold
  return(list(clinic_reporting, date_mat))
}

## Color or size dots by their # from that year
# ctar_data %>%
#   group_by(date_de_consultation = year(date_de_consultation), id_ctar) %>%
#   summarize(tot = n()) -> year_by_clinic

clinic_reporting <- get.days(threshold = 10)[[1]]
# clinic_reporting %>%
#   mutate(prop10 = prop, prop5 = get.days(threshold = 5)[[1]]$prop, 
#          prop30 = get.days(threshold = 30)[[1]]$prop) %>%
#   left_join(year_by_clinic) %>%
#   ggplot(., aes(x = reorder(ctar, prop10), y = prop10)) + 
#   geom_point() +
#   geom_linerange(aes(ymin = prop30, ymax = prop5)) +
#   facet_wrap(~ date_de_consultation, nrow = 1) +
#   coord_flip() +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
#   xlab("") + 
#   ylab("Proportion of days included") +
#   scale_color_gradient(low = "white", high = "blue")
# 
# date_mat <- as.data.frame(get.days(threshold = 10)[[2]])
# date_mat$date_de_consultation <- doses_wide$date_de_consultation
# names(date_mat[1:(ncol(date_mat)-1)]) <- names(doses_wide[2:(ncol(date_mat))])
# library(reshape2)
# head(melt(date_mat, id = "date_de_consultation"))

ggplot(data = patient_ts, aes(date_de_consultation, ctar)) + 
   geom_tile(aes(fill = n)) +
   scale_fill_continuous(type = "viridis", name = "Patients") +
   xlim(start_date, end_date)

ggplot(clinic_reporting, aes(x = date_de_consultation, y = ctar, fill = prop)) + 
  geom_tile() +
  scale_fill_gradient(low = "white", high = "blue")

```



```{r getting Mora data}
patient_ts %>%
  group_by(ctar) %>%
  mutate(exclude = ifelse(n > mean(n) + 6*sd(n), 1, 0)) -> patient_ts

ggplot(data = drop_na(patient_ts, exclude), aes(x = ctar, y = n, color = as.factor(exclude))) +
  geom_jitter(alpha = 0.5, width = 0.25) +
  coord_flip()

to_exclude <- c(grep("contact", unique(ctar_data$remarque), ignore.case = TRUE, value = TRUE),
                grep("contam", unique(ctar_data$remarque), ignore.case = TRUE, value = TRUE),
                grep("consom", unique(ctar_data$remarque), ignore.case = TRUE, value = TRUE),
                grep("passage", unique(ctar_data$remarque), ignore.case = TRUE, value = TRUE))
to_exclude <- to_exclude[-grep("date", to_exclude, ignore.case = TRUE)]
ctar_data$exclude_bycomment <- 0
ctar_data$exclude_bycomment[ctar_data$remarque %in% to_exclude] <- 1
ctar_data <- mutate(ctar_data, date_de_consultation = ymd(date_de_consultation))
patient_ts %>%
  ungroup() %>%
  select(id_ctar, date_de_consultation, exclude_bydate = exclude) %>%
  right_join(ctar_data) -> ctar_check

(ctar_check %$% length(exclude_bycomment[exclude_bycomment == 1 & exclude_bydate == 1]))/
  length(ctar_check %$% exclude_bycomment[exclude_bycomment == 1])

ctar_check %>%
  group_by(id_ctar) %>%
  summarize(prop_excluded = length(exclude_bydate[exclude_bydate == 1])/n()) %>%
  ggplot(data = ., aes(y = id_ctar, prop_excluded)) +
  geom_point()

```


```{r ttimes + pop at commune level}
## Need: weighted ttimes, pop, ctar_in, exclude
comm_catchments %>%
  left_join(select(mada_communes@data, pop = MDG__201, mdg_cm_, district = mtch_ds)) -> comm_covars
comm_covars <- filter(comm_covars, ttimes_weighted != "Inf")
comm_covars %>% 
  left_join(select(mada_district@data, mdg_dis_co, district)) -> comm_data
comm_data$exclude <- dist_catchments$exclude[match(comm_data$mdg_dis_co, dist_catchments$mdg_dis_co)]

## Data to include
nrow(comm_data) - sum(comm_data$exclude)
```


``` {r check with moramanga}
## ctar for # bites and also checking days to rule out contacts
col_lookup <- read.csv("data/new_colnames.csv", header = TRUE)
ctar <- read.csv("data/moramanga/ctar_up_todate.csv", header = TRUE)
ctar <- match.colnames(ctar, subset(col_lookup, form == "ctar"))

## merge with newest data!

## Set-up time series
tfunc <- function (date, format_date = "%Y-%m-%d", start = "01-09-2016", format_start = "%d-%m-%Y",
                   year1 = 2016, tstep = "month", period = FALSE) {
  get.consec (date, format_date, start, format_start,
              year1, tstep, period) - 8 ## starts in Sep
}

## all dates
ctar %>%
  mutate_at(vars(starts_with("date")), funs(dmy(.))) -> ctar
ctar$month <- tfunc(ctar$date.reported)
## subset to study period (Sep 2016 - to date)
## checks for cases where month is NA!
ctar <- subset(ctar, month > 0 & month < 28)

# ctar district and commune
ctar$commune <- sapply(strsplit(as.character(ctar$patient.location), "\\("), "[", 1)
ctar$commune <- trimws(ctar$commune, which = "right")
ctar$district <- sapply(strsplit(as.character(ctar$patient.location), "\\, "), "[", 2)
ctar$district <- gsub(" \\(District\\)", "", ctar$district)

## change all true/false to 0/1
ctar[ctar == FALSE] <- 0

## Bite data in Moramanga
n_months <- max(ctar$month, na.rm = TRUE)

ctar %>%
  left_join(select(mada_communes@data, commune, mdg_cm_, mtch_ds), 
            by = c("commune" = "commune", "district" = "mtch_ds")) %>%
  filter(consult.type == "Animal bite/exposure") %>%
  group_by(mdg_cm_) %>%
  summarize(bites = round(n()/n_months*12)) %>%
  left_join(comm_covars) %>%
  filter(!is.na(ttimes_weighted), 
         district %in% mada_district@data$district[mada_district@data$CTAR == "Moramanga"]) -> mora_bites

ctar %>%
  group_by(date.reported) %>%
  summarize(n = n()) %>%
  mutate(exclude = ifelse(n > mean(n) + 2*sd(n), 1, 0)) -> mora_ts

ggplot(data = mora_ts, aes(x = "Moramanga", y = n, color = as.factor(exclude))) +
  geom_jitter(alpha = 0.5, width = 0.1) +
  coord_flip()

ctar$exclude <-mora_ts$exclude[match(ctar$date.reported, mora_ts$date.reported)]
table(ctar$consult.type, ctar$exclude)

sdev <- seq(2, 10, by = 1)
check <- data.frame(sd = NA, prop_bites = NA, prop_contacts = NA)
for (i in 1:length(sdev)){
  ctar %>%
    group_by(date.reported) %>%
    summarize(n = n()) %>%
    mutate(exclude = ifelse(n > mean(n) + sdev[i]*sd(n), 1, 0)) -> mora_ts
  ctar$exclude <-mora_ts$exclude[match(ctar$date.reported, mora_ts$date.reported)]
  checksd <- table(ctar$consult.type, ctar$exclude)[, 2]/table(ctar$consult.type)
  check <- rbind(check, c(sdev[i], checksd[1:2]))
}

plot(check$sd, check$prop_contacts, col = "blue", pch = 20, bty = "l", ylim = c(0,1))
points(check$sd, check$prop_bites, col = "red", pch = 20)

```

### Models of bites as a function of travel times

```{r district level exposures}
## districts that are non-reporter and data not available yet CTAR
mada_district@data %>%
select(district = mdg_dis_co, pop = pop2015adj, ttimes_weighted, CTAR, exclude) -> covars

ctar_check %>% 
  filter(exclude_bydate == 0, exclud_bycomment == 0) %>% # exclude contacts
  mutate_at(vars(starts_with("date")), funs(ymd(.))) %>% # format dates
  group_by(year = year(date_de_consultation), district, id_ctar) %>% # group by year and district
  summarize(n = n()) %>% # get count of exposures
  left_join(covars) %>% # add pop + ttimes + reporting
  left_join(clinic_reporting) %>%   # add reporting by clinic + year
  filter(exclude == 0, prop > 0.25) %>%
  group_by(district, year) %>%
  mutate(bites = n/prop) %>%
  summarise(bites = sum(bites, na.rm = TRUE)) %>%
  group_by(district) %>%
  summarise(bites = mean(bites)) %>%
  left_join(covars) -> exps_check

```


### Model results Moramanga and District
```{r likelihood of observing moramanga data @ commune level}

get.likelihood.mora <- function(commune, params) {
  # To test likelihood function
  # commune = comm_data
  # district = exps_dist
  # params = c(1e-6, 0.1)
  B_ttimes <- params[1]
  B_0 <- params[2]
  commune %>%
      mutate(bites_mu = exp(B_ttimes*ttimes_weighted + B_0)*pop) -> sum_bites
  return(-sum(dpois(round(sum_bites$bites), lambda = sum_bites$bites_mu, log = TRUE)))
}

get.likelihood <- function(commune, district, params) {
  # To test likelihood function
  # commune = comm_data
  # district = exps_dist
  # params = c(1e-6, 0.1)
  B_ttimes <- params[1]
  B_0 <- params[2]
  commune %>%
      filter(exclude == 0) %>%
      group_by(mdg_dis_co) %>%
      mutate(bites = exp(B_ttimes*ttimes_weighted + B_0)*pop) %>%
      summarize(sum_bites = sum(bites)) %>%
      left_join(select(district, district, dist_bites = bites), 
                by = c("mdg_dis_co" = "district")) %>%
      filter(!is.na(dist_bites)) -> sum_bites
  return(-sum(dpois(round(sum_bites$dist_bites), lambda = sum_bites$sum_bites, log = TRUE)))
}
```



```{r likelihood of observing district data given sum of communes in district}
mora_pars <- optim(par = c(1e-6, 0.1), get.likelihood.mora, commune = mora_bites, hessian = TRUE)
B_ttimes_mora <- mora_pars$par[1]
B_0_mora <- mora_pars$par[2]

## Getting standard errors
# this is how you can get standard errors from optim, but inverting the hessian
se <- sqrt(abs(diag(solve(mora_pars$hessian)))); se

#simulate "standard errors"
varcov <- solve(mora_pars$hessian)
library(MASS)

ntest = 1000; testpar = mora_pars$par; sim.par = matrix(NA, ntest, 2)
for (k in 1:ntest) {
  sim.par[k, ] = mvrnorm(1, mora_pars$par, varcov)
}

B_ttimes_mora_CI <- quantile(sim.par[,1],c(0.025,0.975))
B_0_mora_CI <- quantile(sim.par[,2],c(0.025,0.975))

mada_pars <- optim(par = c(1e-6, 0.1), get.likelihood, district = exps_check,
             commune = comm_data, hessian = TRUE)
B_ttimes_mada <- mada_pars$par[1]
B_0_mada <- mada_pars$par[2]

## Getting standard errors
# this is how you can get standard errors from optim, but inverting the hessian
se <- sqrt(abs(diag(solve(mada_pars$hessian)))); se

#simulate "standard errors"
varcov <- solve(mada_pars$hessian)
library(MASS)

ntest = 1000; testpar = mada_pars$par; sim.par = matrix(NA, ntest, 2)
for (k in 1:ntest) {
  sim.par[k, ] = mvrnorm(1, mada_pars$par, varcov)
}

B_ttimes_mada_CI <- quantile(sim.par[,1],c(0.025,0.975))
B_0_mada_CI <- quantile(sim.par[,2],c(0.025,0.975))
par_df <- data.frame(values = c(B_ttimes_mada, B_ttimes_mora, B_0_mada, B_0_mora),  
                     par = c("B_ttimes", "B_ttimes", "intercept", "intercept"),
                     upper_CI = c(CI = B_ttimes_mada_CI[2], B_ttimes_mora_CI[2],
                                  B_0_mada_CI[2], B_0_mora_CI[2]),
                     lower_CI = c(CI = B_ttimes_mada_CI[1], B_ttimes_mora_CI[1],
                                  CI = B_0_mada_CI[1], B_0_mora_CI[1]),
                     model = c("Mada", "Mora", "Mada", "Mora"))

ggplot(data = filter(par_df, par == "B_ttimes"), aes(x = model, y = values)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower_CI, ymax = upper_CI))
ggplot(data = filter(par_df, par == "intercept"), aes(x = model, y = values)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower_CI, ymax = upper_CI))

ttimes_plot <- seq(0, 15*60, by = 30)
pop_plot <- 1e5

## Mada
plot(ttimes_plot/60, exp(B_ttimes_mada*ttimes_plot + B_0_mada)*pop_plot, type = "l", bty = "n",
     ylab = "Mean bites per 100k", xlab = "Travel times (hrs)", col = "darkblue")
points(ttimes_plot/60, exp(B_ttimes_mada_CI[1]*ttimes_plot + B_0_mada_CI[1])*pop_plot, type = "l", 
       lty = 2, col = "lightblue")
points(ttimes_plot/60, exp(B_ttimes_mada_CI[2]*ttimes_plot + B_0_mada_CI[2])*pop_plot, type = "l", 
       lty = 2, col = "lightblue")


points(ttimes_plot/60, exp(B_ttimes_mada*ttimes_plot + B_0_mada)*pop_plot, pch = 20, col = "darkblue")
points(ttimes_plot/60, exp(B_ttimes_mada_CI[1]*ttimes_plot + B_0_mada_CI[1])*pop_plot, type = "l", 
       lty = 2, col = "lightblue")
points(ttimes_plot/60, exp(B_ttimes_mada_CI[2]*ttimes_plot + B_0_mada_CI[2])*pop_plot, type = "l", 
       lty = 2, col = "lightblue")

points(ttimes_plot/60, exp(B_ttimes_mora*ttimes_plot + B_0_mora)*pop_plot, type = "l", col = "purple")
points(ttimes_plot/60, exp(B_ttimes_mora_CI[1]*ttimes_plot + B_0_mora_CI[1])*pop_plot, type = "l", 
       lty = 2, col = "magenta")
points(ttimes_plot/60, exp(B_ttimes_mora_CI[2]*ttimes_plot + B_0_mora_CI[2])*pop_plot, type = "l",
       lty = 2, col = "magenta")
legend("topright", c("Mada", "Mora", "CIs"), col = c("darkblue", "purple", "grey"), lty = c(1, 1, 2), 
       bty = "n")

## Data vs. predicted
comm_data %>%
      filter(exclude == 0) %>%
      group_by(mdg_dis_co) %>%
      mutate(bites = exp(B_ttimes*ttimes_weighted + B_0)*pop) %>%
      summarize(sum_bites = sum(bites)) %>%
      left_join(select(exps_dist, district, dist_bites = bites), 
                by = c("mdg_dis_co" = "district")) %>%
      filter(!is.na(dist_bites)) -> sum_bites
sum_bites %$% plot(dist_bites, sum_bites, bty = "n", pch = 20, 
                   xlab = "Observed mean bites at district (Mada)", ylab = "Model predicted mean bites")
abline(a = 0, b = 1, col = "blue", lty = 2)

bites %>%
      mutate(bites_mu = exp(B_ttimes*ttimes_weighted + B_0)*pop) -> check
plot(check$bites, check$bites_mu, bty = "n", pch = 20, 
      xlab = "Observed mean bites at commune (Moramanga)", ylab = "Model predicted mean bites")
abline(a = 0, b = 1, col = "blue", lty = 2)

```


### Adding dist with ctar vs. without

```{r checking impact of ctar in}
exps_dist$ctar_in <- dist_catchments$ctar_in_dist[match(exps_dist$district,
                                                        dist_catchments$mdg_dis_co)]
ggplot() + 
  geom_boxplot(data = exps_dist, aes(x = as.factor(ctar_in), y = bites/pop))

comm_data$ctar_in_comm <- comm_catchments$ctar_in_comm[match(comm_data$mdg_cm_,
                                                        comm_catchments$mdg_cm_)]
comm_data$ctar_in_dist <- dist_catchments$ctar_in_dist[match(comm_data$mdg_dis_co,
                                                        dist_catchments$mdg_dis_co)]

get.likelihood <- function(commune,district, params) {
  # To test likelihood function
  # commune = comm_data
  # district = exps_dist
  # params = c(1e-6, 0.1)
  B_ttimes <- params[1]
  B_in <- params[2]
  B_0 <- params[3]

  commune %>%
      filter(exclude == 0) %>%
      group_by(mdg_dis_co) %>%
      mutate(bites = exp(B_ttimes*ttimes_weighted + B_in*ctar_in_dist + B_0)*pop) %>%
      summarize(sum_bites = sum(bites)) %>%
      left_join(select(district, district, dist_bites = bites), 
                by = c("mdg_dis_co" = "district")) %>%
      filter(!is.na(dist_bites)) -> sum_bites
  return(-sum(dpois(round(sum_bites$dist_bites), lambda = sum_bites$sum_bites, log = TRUE)))
}

tmp <- optim(par = c(1e-6, 1e-6, 0.1), get.likelihood, district = exps_dist,
             commune = comm_data, hessian = TRUE)
params <- tmp$par
B_ttimes <- params[1]
B_in <- params[2]
B_0 <- params[3]

comm_data %>%
      filter(exclude == 0) %>%
      group_by(mdg_dis_co) %>%
      mutate(bites = exp(B_ttimes*ttimes_weighted + B_in*ctar_in_comm + B_0)*pop) -> check
plot(check$ttimes_weighted, check$bites)

comm_data %>%
      filter(exclude == 0) %>%
      group_by(mdg_dis_co) %>%
      mutate(bites = exp(B_ttimes*ttimes_weighted + + B_in*ctar_in_comm + B_0)*pop) %>%
      summarize(sum_bites = sum(bites)) %>%
      left_join(select(exps_dist, district, dist_bites = bites), 
                by = c("mdg_dis_co" = "district")) %>%
      filter(!is.na(dist_bites)) -> sum_bites
sum_bites %$% plot(dist_bites, sum_bites)
abline(a = 0, b = 1, col = "blue", lty = 2)


```


```{burden by incidence}

## Mada
plot(ttimes_plot/60, exp(B_ttimes_mada*ttimes_plot + B_0_mada)*pop_plot, type = "l", bty = "n",
     ylab = "Mean bites per 100k", xlab = "Travel times (hrs)", col = "darkblue")
points(ttimes_plot/60, exp(B_ttimes_mada_CI[1]*ttimes_plot + B_0_mada_CI[1])*pop_plot, type = "l", 
       lty = 2, col = "lightblue")
points(ttimes_plot/60, exp(B_ttimes_mada_CI[2]*ttimes_plot + B_0_mada_CI[2])*pop_plot, type = "l", 
       lty = 2, col = "lightblue")
abline(h = pop_plot/25*0.01*0.39, col = "darkred") # low rabies incidence
abline(h = pop_plot/5*0.01*0.39, col = "red") # high rabies incidence

bites_100k <- exp(B_ttimes_mada*ttimes_plot + B_0_mada)*pop_plot
R_100k_min <- pop_plot/25*0.01*0.39
R_100k_max <- pop_plot/5*0.01*0.39

plot(ttimes_plot/60, 
     bites_100k*(ifelse(bites_100k*0.6 > R_100k_min, R_100k_min/bites_100k, 0.6))/R_100k_min, 
     type = "l", bty = "n",
     ylab = "Reporting", xlab = "Travel times (hrs)", col = "darkred")
lines(ttimes_plot/60, 
     bites_100k*(ifelse(bites_100k*0.2 > R_100k_max, R_100k_max/bites_100k, 0.2))/R_100k_max,
     col = "red")

ests_lower <- predict(mod_lower, newdata = as.data.frame(ttimes_plot), 
                                type = "response")
ests_upper <- predict(mod_upper, newdata = as.data.frame(ttimes_plot), 
                                type = "response")

plot(ttimes_plot/60, 
     bites_100k*(ifelse(bites_100k*ests_upper > R_100k_min, R_100k_min/bites_100k,
                        ests_upper))/R_100k_min, 
     type = "l", bty = "n",
     ylab = "Reporting", xlab = "Travel times (hrs)", col = "darkred")
lines(ttimes_plot/60, 
     bites_100k*(ifelse(bites_100k*ests_lower > R_100k_max, R_100k_max/bites_100k,
                        ests_lower))/R_100k_max,
     col = "red")

```

### Estimation of burden and reporting 

```{r functions for getting burden}
get.burden <- function(covs = covars, pop = covars$pop, model = mod,
                       p_rabid = 0.5, rho_max = 0.9,
                       max_HDR = 25, min_HDR = 5, 
                       dog_rabies_inc = 0.01, human_exp_rate = 0.39, 
                       prob_death = 0.16, nsims = 1000) {
  store_deaths <- store_p_rabid <- matrix(NA, nrow(covs), nsims)

  for (i in 1:nsims) {
    mean_bites <- predict(model, type = "response", newdata = covs)
    bites <- rpois(length(mean_bites), mean_bites)
    mean_exps <- runif(length(pop), min = human_exp_rate*dog_rabies_inc*pop/max_HDR, 
                    max = human_exp_rate*dog_rabies_inc*pop/min_HDR)
    rabid_exps <- rpois(length(mean_bites), mean_exps)
    p_rabid_t <- ifelse(bites*p_rabid/rabid_exps > rho_max, (rho_max*rabid_exps)/bites, p_rabid)
    deaths <- rbinom(length(mean_bites), rabid_exps - round(bites*p_rabid_t), prob_death)
    store_deaths[, i] <- deaths
    store_p_rabid[, i] <- p_rabid_t
  }
  
  results <- cbind(get.meanCI(store_deaths), get.meanCI(store_p_rabid))
  colnames(results) <- c("deaths_mean", "deaths_lowerCI", "deaths_upperCI",
                         "p_rabid_mean", "p_rabid_lowerCI", "p_rabid_upperCI")
  results <- cbind(covars, results)
  return(results)
}

get.burden.flat <- function(covs = covars, pop = covars$pop, model = mod,
                       p_rabid = 0.5, rho = 0.85, 
                       prob_death = 0.16, nsims = 1000) {
  store_deaths <- matrix(NA, nrow(covs), nsims)

  for (i in 1:nsims) {
    mean_bites <- predict(model, type = "response", newdata = covs)
    bites <- rpois(length(mean_bites), mean_bites)
    unreported <- bites*p_rabid/rho - bites*p_rabid
    deaths <- rbinom(length(mean_bites), round(unreported), prob_death)
    store_deaths[, i] <- deaths
  }
  results <- get.meanCI(store_deaths)
  colnames(results) <- c("deaths_mean", "death_lowerCI", "deaths_upperCI")
  results <- cbind(covars, results)
  return(results)
}

get.meanCI <- function(matrix_by_row, nboots = 1000) {
    mean_rows <- apply(matrix_by_row, 1, mean)
    CIs <- apply(matrix_by_row, 1, boots, nboot = nboots)
    lower <- unlist(CIs)[1, ] 
    upper <- unlist(CIs)[2, ]
    summ <- cbind(mean_rows, lower, upper)
    colnames(summ) <- c("mean", "lower", "upper")
    return(summ)
}

boots <- function(x, nboot = 1000){
  n = length(x)
  xbar = mean(x)
  tmpdata = sample(x,n*nboot, replace=TRUE)
  bootstrapsample = matrix(tmpdata, nrow=n, ncol=nboot)
  bsmeans = colMeans(bootstrapsample)
  deltastar = bsmeans - xbar
  d = quantile(deltastar, c(0.05, 0.95))
  ## Get from quantiles
  ci = xbar - c(d[2], d[1])

  ## Sorted CIs
  sorteddeltastar = sort(deltastar)
  # # Look at the sorted results
  # Find the .05 and .95 critical values of deltastar
  d9alt = sorteddeltastar[5]
  d1alt = sorteddeltastar[950]
  # Find and print the 80% confidence interval for the mean
  ciAlt = xbar - c(d1alt,d9alt)
  # cat(’Alternative confidence interval: ’,ciAlt, ’\n’)
  return(ci)
}

```

```{r}
results_pr20 <- get.burden(p_rabid = 0.20, rho_max = 0.98)
results_pr60 <- get.burden(p_rabid = 0.60,rho_max = 0.98)

plot(results_pr20$ttimes_weighted/60, results_pr20$deaths_mean, col = "darkred", pch = 20)
points(results_pr60$ttimes_weighted/60, results_pr60$deaths_mean, col = "red", pch = 20)
segments(x0 = results_pr60$ttimes_weighted/60, x1 = results_pr60$ttimes_weighted/60,
         y0 = results_pr60$deaths_mean, y1 = results_pr20$deaths_mean, col = "red")

results_base <- get.burden(p_rabid = 0.40, rho_max = 0.98)

p_rabid_check <- seq(0.10, 0.80, by = 0.1)
check <- data.frame(p_rabid = 0, district = NA, ttimes = 0, deaths_mean = 0)
for (i in 1:length(p_rabid_check)) {
  print(paste(i, "/", length(p_rabid_check)))
  results <- get.burden(p_rabid = p_rabid_check[i], rho_max = 0.98)
  p_rabid <- rep(p_rabid_check[i], length(results$deaths_mean))
  check_one <- cbind(p_rabid, results$district, results$ttimes, results$deaths_mean)
  colnames(check_one) <- c("p_rabid", "district", "ttimes", "deaths_mean")
  check <- rbind(check, check_one)
}

check <- check[-1, ]
check$base <- rep(results_base$deaths_mean, length(p_rabid_check))
ggplot(data = check, aes(x = as.factor(p_rabid), y = as.factor(reorder(district, ttimes)))) + 
  geom_tile(aes(fill = deaths_mean/base)) +
  scale_fill_gradient2(high = "red", mid = "white", low = "blue", midpoint = 1, name = "Mean deaths") 

rho_max_check <- seq(0.90, 0.99, by = 0.01)
check <- data.frame(rho_max = 0, district = NA, ttimes = 0, deaths_mean = 0)
for (i in 1:length(rho_max_check)) {
  print(paste(i, "/", length(rho_max_check)))
  results <- get.burden(p_rabid = 0.4, rho_max = rho_max_check[i])
  rho_max <- rep(rho_max_check[i], length(results$deaths_mean))
  check_one <- cbind(rho_max, results$district, results$ttimes, results$deaths_mean)
  colnames(check_one) <- c("rho_max", "district", "ttimes", "deaths_mean")
  check <- rbind(check, check_one)
}

check <- check[-1, ]
check$base <- rep(results_base$deaths_mean, length(rho_max_check))
check %>%
  group_by(district) %>%
  mutate(min = min(deaths_mean), max = max(deaths_mean), 
         standardized = (deaths_mean - base)/(max - base)) -> check

ggplot(data = check, aes(x = as.factor(p_rabid), y = as.factor(reorder(district, ttimes)))) + 
  geom_tile(aes(fill = standardized)) +
  scale_fill_gradient2(high = "red", mid = "white", low = "blue", midpoint = 0, name = "Mean deaths") 
ggplot(data = check, aes(x = as.factor(p_rabid), y = as.factor(reorder(district, ttimes)))) + 
  geom_tile(aes(fill = deaths_mean)) +
  scale_fill_gradient2(high = "red", mid = "white", low = "blue", midpoint = 0, name = "Mean deaths") 

mada_dist_gg <- tidy(mada_district, region = "mdg_dis_co")
mada_dist_gg <- left_join(mada_dist_gg, exps_cleaned, by = c("id" = "district"))

ggplot() +
  geom_polygon(data = mada_dist_gg, aes(x = long, y = lat, group = group, fill = CTAR)) +
  scale_colour_viridis_d()

```


### Sensitivity Analyses

### Estimating the impact of expanding PEP Access

<br>

## Discussion

### Key findings

### Strengths and Limitations
- Not accounting for clinic functioning
- Not explicit data on reporting or rabies incidence!
- Other factors that drive reporting
- While we use a signficiant amount of assumptions, these are the same set of assumptions that underlie current rabies estimates, we're just improving them a bit... and ours are a likely a little bit more in line with reality @ a sub-national level

### Broader context

### Conclusions
